\documentclass{amsart}
\usepackage{amsmath, amssymb, amscd}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.25in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}

\begin{document}

\section*{Math 4324\  Second Compactness Assignment} 

\subsection*{Hand in Friday, March 29.} 

\vspace{.15in}
In this assignment I'll hyphenate ``$\omega$-accumulation-point-compact" to make it clearer that the full expression is one adjective describing a space. 


\vspace{.15in}
\noindent
\subsection*{1.}  Show that, if the topological space $X$ is compact, then every infinite subset of $X$ has an $\omega$-accumulation point in $X$, i.e. that every compact space is $\omega$-accumulation-point-compact. {\bfseries Hint.} If $Y$ is a subset that has no $\omega$-accumulation point, choose a neighborhood of each point of $X$ that helps you show that $Y$ must be finite. 

\begin{proof}
    Let $X$ be a compact topological space and $Y\subset X$ be a set with no $\omega-$accumulation point. Then for each $y \in Y$ we have a neighborhood $y\in U_y$ but $\mathcal U_y=Y\cap U_y$ is finite. Then we create the open cover $\{\mathcal U_y\}_{y\in Y}$ but as $X$ is compact we have a finite subcover $\{\mathcal U_{y_1},\ldots,\mathcal U_{y_n}\}$. But then $Y= \bigcup_{i=1}^n \mathcal U_{y_i}$ and so $Y$ is finite. This shows that any set that does not have an $\omega-$accumulation point is finite hence any infinite set must have an $\omega-$accumulation point.
\end{proof}


\vspace{.15in}
\noindent
\subsection*{2.} Show that, if a topological space $X$ is sequentially compact (every sequence in $X$ has a subsequence converging to a point in $X$), then $X$ is $\omega$-accumulation-point-compact. 

\begin{proof}
    Assume that $X$ is a sequentially compact. Then take an arbitrary infinite set $A\subset X$ then create an arbitrary sequence $\{a_n\}_{n\in \mathbb{N}}$ in $A$. 
    
    
    Then as $X$ is sequentially compact we have a subsequence $\{a_{n_k}\}_{k\in \mathbb{N}}$ for some $a\in A$ we have any neighborhood $U$ of $a$ which contains all points of the subsequence after some $n_k\in \mathbb{N}$. 
    Hence $U$ contains an infinite number of elements of $A$ we have that $a$ is an $\omega-$accumulation point of $A$. As $A$ was an arbitrary infinite set we have that $X$ is $\omega-$accumulation-point-compact.
\end{proof}

\vspace{.15in}
\noindent
\subsection*{3.} Show that, if a topological space $X$ is $\omega$-accumulation-point-compact and first countable, then $X$ is sequentially compact. {\bfseries Hint.} For an arbitrary sequence, show that, if no single value appears with infinitely many indices, then the sequence must have infinitely many distinct values. Use first countable to show that an $\omega$-accumulation point is a limit of a subsequence. 


\begin{proof}
    Let $X$ be a first countable $\omega-$accumulation-point-compact space. Then take an arbitrary sequence $\{x_n\}_{n\in \mathbb{N}}$ in $X$. In the event that some value in our sequence occurs an infinite number of times we can just create the constant subsequence and we are done. So we have that each value in our sequence occurs finitely many times hence our sequence contains infinitely many distinct points. 

    Then as $\{x_n\}_{n\in \mathbb{N}}$ is an infinite subset of $X$ we have for some $x \in \{x_n\}_{n\in \mathbb{N}}$ is an $\omega-$accumulation point. 
    
    Now as $X$ is first countable take the countable basis $\{B_n\}_{n\in \mathbb{N}}$ of $x$ each of which intersects $\{x_n\}_{n\in \mathbb{N}}$ an infinite number of times. Then create the subsequence $x_{n_k}\in \bigcap_{i=1}^{n_k} (B_i ) \cap \{x_n\}_{n\in \mathbb{N}}$ where for each $n_k<n_{k+1}$ such an element is always guaranteed to exist as the intersection is infinite. 
    
    Lastly to show that this subsequence converges and has limit point $x$. Let $U_x$ be an arbitrary neighborhood of $x$ then we have as $X$ is first countable that for some $B\in \{B_n\}_{n\in \mathbb N}$  we have $B\subset U_x$ and for some $n_K\in \mathbb{N}$ we have that for all $n_k \geq n_K$ that $x_{n_k} \in B\subset U_x$ hence we have that this subsequence converges to $x$ and so $X$ is sequentially compact.

    
\end{proof}

\vspace{.15in}
\begin{definition} A topological space $X$ is called {\it second countable} if $X$'s topology has a countable basis. \end{definition}


\vspace{.15in}
\noindent
\subsection*{4.} 

\noindent
{\bfseries a.} Show that, if a topological space $X$ is second countable, then every open cover of $X$ has a countable subcover.

\begin{proof}
    Suppose that $X$ is second countable $A\subset X$ and $\{U_a\}_{a\in B  }$ is an arbitrary open cover of $A$. Then we have a countable basis covering of $A$ where $\{B_a\}_{a \in A}$ where 
\end{proof}

\vspace{.1in}
\noindent
{\bfseries b.} Show that every second countable, $\omega$-accumulation-point-compact space $X$ is compact. {\bfseries Hint.} For a countable open cover $\{ U_j\}$ with no finite subcover, construct an infinite subset $Y$ of $X$ with the property that each $U_j$ contains at most finitely many elements of $Y$. 

\vspace{.15in}
\begin{definition} In a metric space $(X,d)$, the {\it diameter} of a nonempty bounded subset $Y$ is the supremum (least upper bound) of $\{ d(y_1, y_2) : y_1, y_2 \in Y\}$. \end{definition}

\vspace{.15in}
\begin{definition} Let $\{ U_{\alpha}\} _{\alpha \in A}$ be an open cover of a metric space $X$. A {\it Lebesgue number} for this cover is a number $\delta$ satisfying: for every subset $Y$ of $X$ with diameter less than $\delta$, there is a $U_{\alpha}$ for which $Y\subset U_{\alpha}$. \end{definition}

\vspace{.15in}
Problem 5 is an {\bf extra credit} problem, again with no penalty for students who do not do problem 5, as a way of making sure the assignment is of reasonable length.
 
\vspace{.15in}
\noindent
\subsection*{5.} The results in this problem apply to any metrizable topological space, that is to any space whose topology comes from a metric, regardless of which metric is used, as long as the topology the metric generates is the space's given topology. So that no one will worry about the implications of a choice of metric, I'll phrase the problem in terms of a metric space, i.e. I'll choose at the beginning a specific metric that defines the topology.  

\vspace{.1in}
\noindent
{\bfseries a.} Show that, if $X$ is an $\omega$-accumulation-point-compact metric space, then every open cover of $X$ has a Lebesgue number. {\bfseries Hint.} Consider the implications of having a sequence of subsets $(Y_n)$, satisfying, for each $n$, the diameter of $Y_n$ is less than $1/n$ and, for each $n$, no open set from the cover contains $Y_n$.  

\vspace{.1in}
\noindent
{\bfseries b.} Show that, if $X$ is an $\omega$-accumulation-point-compact metric space, then, for each $\epsilon > 0$, there is a finite collection of open balls of radius $\epsilon$ that covers $X$. {\bfseries Hint.} If there is some $\epsilon > 0$ for which no finite collection of open balls of radius $\epsilon$ that covers $X$, exhibit an infinite set with no $\omega$-accumulation point.

\vspace{.1in}
\noindent
{\bfseries c.} Show that, if $X$ is an $\omega$-accumulation-point-compact metric space, then $X$ is compact. {\bfseries Hint.} Use parts a. and b. 


 
\end{document}