\documentclass{amsart}
\usepackage{amsmath, amssymb, amscd}

\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9in}
\setlength{\topmargin}{-.25in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}
\newtheorem{notation}[theorem]{Notation}

\begin{document}

\section*{Math 4324\  Applications of the winding number} 

\subsection*{Hand in Monday, April 26.} 

\vspace{.25in}

\vspace{.25in}
\noindent
\subsection*{1.} Let $D$ be  a disk with boundary circle $C$, and let $f: D\to \mathbb{R}^2$ be a continuous map. Suppose $P$ is a point in $\mathbb{R}^2$ with $P\not \in f(C)$, and the winding number of the restriction $f|_C$ of $f$ to $C$ around $P$ is not zero. Show that there is some point $Q$ in $D$ such that $f(Q) = P$.

\begin{proof}

    Suppose that $f:D\to \mathbb{R}^2$ where $D$ has boundary circle $C$. Suppose that $P\in \mathbb{R}^2$ with $P\not \in f(D)$. Choose any point $Q\in C$ and and create the function $g:C\times [0,1]\to D$ by the equation $g(\vec v, t)=\vec v(1-t)+t \vec c$ this is a continuous function as it is a weighted sum of continuous functions. Now we have that $h:C\times [0,1]\to \mathbb{R}^2$ defined by the equation $h(\vec v,t)=f(g(\vec v,t))$ is a homotopy between $f|_C$ and the constant curve at $\vec c$. This is shown to be a homotopy as $h(\vec v,0)=f(g(\vec v,0))= f(\vec v)$ and $h(\vec v,1)=f(g(\vec v,1))=f(\vec c)$. We have that $h$ is continuous as it is the composition of two continuous functions. As the constant curve at $\vec c$ has winding number $0$ we get that $W(f|_C,P)=0$. This proves the contrapositive of the statement.
    
\end{proof}

\vspace{.15in}
\noindent
\subsection*{2.}

If $f:C\to C$  is a continuous mapping with no fixed point, show that degree of $f$ must be $1$. 
\begin{proof}
    It suffices to assume that $C$ is centered at the origin as we could create a homeomorphism that shifts it. Now assume that $f:C\to C$ is a continuous map with no fixed points. Then consider the homotopy $h:[0,1]\times C\to C$ defined by the equation $h(t,\vec c)=(1-t)f(\vec c)+t(-\vec c)$ this is continuous as it is two weighted continuous functions. As no point is fixed we get the line that intersects $f(\vec c)$ and $g(\vec c)=-\vec c$ does not intersect $\vec 0$. Lastly as $h(0,\vec c)=f(\vec c)$ and $h(1,\vec c)=-\vec c$ we get that $h$ is a homotopy. 
    
    On problem 2b on last weeks homework we showed that $\gamma(x,y)=(-x,-y)$ on the unit circle to the unit circle has degree 1. We have that $g$ is homotopic to $\gamma$ (changing radius of circle) so we get that $g$ has degree 1. As $f$ is homotopic to $g$ we get that $f$ has degree 1.

\end{proof}


\vspace{.15in}
\noindent
\subsection*{3.}
If $f:D^2\to \mathbb{R}^2$ is continuous and $f(P)\cdot P\not = 0$ for all $P$ in $S^1$, show that there is some $Q\in D^2$ with $f(Q)=0$.

\begin{proof}
    Assume that $f:D^2\to \mathbb{R}^2$ is continuous and $f(P)\cdot P\not =0$ for all $P\in S^1$. Then as for all $\vec c\in C$ we can not have the line that intersects $f(\vec c)$ and $g(\vec c)$ intersects $\vec 0$ which follows from the fact that $f(\vec v)\cdot \vec v\not =0 $ and $g(\vec v)\cdot \vec 0=0$ hence we get by the Dog-on-a-Leash theorem that $W(f|_{S^1},\vec 0)=W(g,\vec 0)$. As $W(g,\vec 0)$ has a non zero winding number then we get by problem 1 that there exists a point $Q\in D^2$ such that $f(Q)=0$.

\end{proof}


\vspace{.15in}
\noindent
\subsection*{4.} Show that if $f:C\to C^\prime$ is a map between circles such that $f(P^\star)=f(P)$ for all $P$, then the degree of $f$ is even.

\begin{proof}
We have $C,C^\prime$ are both homeomorphic to $S^1$. So let $f:S^1\to S^1$ such that $f(P^\star)=f(P)$ for all $P$. 

Let $\gamma_1(\theta)=f(\cos \theta ,\sin \theta )$ for $0\leq \theta \leq \pi $ and $\gamma_2(\theta)=f(\cos \theta +\pi , \sin \theta+\pi)$ for $\pi \leq \theta \leq 2\pi$. 

As $f(\cos \theta,\sin \theta )=f(\cos \theta+\pi,\sin \theta+\pi)$ we get both $\gamma_1,\gamma_2$ are closed path hence $W(\gamma_1,\vec 0),W(\gamma_2,\vec 0)$ both have integer winding numbers. As $\gamma_1=\gamma_2$ we get $W(\gamma_1,\vec 0)=W(\gamma_2,\vec 0)=n$ for some $n\in \mathbb Z$. Hence we have $W(f,\vec 0)=W(\gamma_1,\vec 0)+W(\gamma_2,\vec 0)=2n$. Hence the degree of $f$ is even.

\end{proof}

\vspace{.15in}
\noindent
\subsection*{5.}
If $f$ and $g$ are continuous real-valued functions on a sphere $S$ such that that $f(P^\star)=-f(P)$ and $g(P^\star)=-g(P)$ for all $P$, show that $f$ and $g$ must have a common zero on the sphere. 

\begin{proof}
    Create the function $h:S\to \mathbb{R}^2$ where $h(\vec x)=(f(\vec x),g(\vec x))$ this is continuous by Munkres Theorem 18.4. Then we have by Borsuk-Ulam theorem that there exists a point such that $h(\vec x)=h(\vec x^\star)$ this implies $(f(\vec x),g(\vec x))=(-f(\vec x),-g(\vec x))$ hence we get $f(\vec x)= 0$ and $g(\vec x)=0$.
\end{proof}

 
\end{document}